Palaeodistribution of crop progenitors and wild foods in the Southern Levant

### 2023-08-04 analysis/analysis.Rmd
Tune RF model
Use a classification forest, probability (relative likelihood) output, down-sampling to deal with class imbalance

```{r psabsences, include=FALSE}
n_background <- 1000

occ |>
  mutate(present = TRUE) |>
  bind_rows(background_sample(s_levant_land, 
                              n_background,
                              coord_x = "decimalLongitude",
                              coord_y = "decimalLatitude", 
                              taxon = "Hordeum spontaneum", 
                              present = FALSE)) ->
  occ

# Weights
#occ %>% 
#  group_by(taxon, occurrence) %>% 
#  mutate(weight = 1 + (n_psab / n() - 1)) ->
#  occ
```

@@

```{r model}
# TODO: Where do the missing value(s) actually come from?
occ |>
  as_tibble() |>
  select(present, starts_with("bio")) |>
  drop_na() ->
  rf_data

rand_forest(mode = "regression") |>
  fit(as.numeric(present) ~ ., data = rf_data) ->
  rf_fit
```

@@

```{r predict}
rf_pred <- predict(climate$cur, rf_fit)
```

>>>>

```{r psabsences, include=FALSE}
n_background <- 10000

occ |>
  mutate(present = TRUE) |>
  bind_rows(background_sample(s_levant_land, 
                              n_background,
                              coord_x = "decimalLongitude",
                              coord_y = "decimalLatitude", 
                              taxon = "Hordeum spontaneum", 
                              present = FALSE)) ->
  occ
```

@@

There are many methods for ENM.
Random Forest is one of the best performing, providing it is appropriately parameterised to account for the class imbalance between presence and background samples (balanced random forest) [@ValaviEtAl2021a; @ValaviEtAl2021b].

We used the R package ranger [@WrightZiegler2017].

```{r model}
# TODO: Where do the missing value(s) actually come from?
# Parameters adapted from the example of a down-sampled RF model given in
#   https://rvalavi.github.io/SDMwithRFs/#model-fitting-with-ranger-package
occ |>
  as_tibble() |>
  select(present, starts_with("bio")) |>
  drop_na() ->
  rf_data

# Model
# TODO: tune?
rand_forest(mode = "classification", trees = 1000) |>
  set_engine(
    "ranger",
    probability = TRUE,
    sample.fraction = c(2250 / n_background, 1), # TODO: parameterise
  ) -> 
  rf_model

# Fit
rf_fit <- fit(rf_model, factor(present) ~ ., data = rf_data)
```

@@

```{r predict}
# TODO: predict on all periods (maybe combine stars and collapse into dimension first?)
rf_pred_cur <- predict(climate$cur, rf_fit)
rf_pred_eh <- predict(climate$EH, rf_fit)
```
